Automated Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{NSF Grant ???} }

### 2021-03-31 paper/paper_DEB.tex

%\documentclass[acmsmall,screen,review]{acmart}
%\documentclass[acmconf,screen,review]{acmart}
\documentclass[screen,review]{acmart}
\usepackage{todonotes}
\usepackage{comment}
\usepackage{wrapfig}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2021}
\acmYear{2021}
\acmDOI{}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[JCDL '21]{JCDL '21: Joint Conference on
  Digital Libraries, September 27--30, 2021, Urbana--Champaign, Illinois, USA}
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
  %June 03--05, 2018, Woodstock, NY}
%\acmPrice{15.00}
%\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Automated Metadata Generation for Fish Specimen Image Collections}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Joel Pepper}
\email{jcp353@drexel.edu}
\orcid{0002-1601-8729}
\author{Jane Greenberg}
\orcid{0001-7819-5360}
\affiliation{%
  \institution{Drexel University}
  \streetaddress{3675 Market St}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
  \postcode{19104}
}
\author{Yasin Baki\c{s}}
\orcid{0001-6144-9440}
\author{Henry Bart Jr.}
\orcid{0002-5662-9444}
\affiliation{
  \institution{Tulane University}
  \city{New Orleans}
  \state{Louisiana}
  \country{USA}
}
\author{David Breen}
\orcid{0002-1376-5008}
\affiliation{%
  \institution{Drexel University}
  \streetaddress{3675 Market St}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
  \postcode{19104}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Pepper, Greenberg, Baki\c{s}, Bart and Breen}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
%%    \todo{Just throwing the Hank conference abstract in here for us to work off of and pare down. Also, we need to decide what to do for the ``teaser image''.} 
%    Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been collected worldwide by museums and other research institutions. The scanning of these specimen collections and the placement of the resulting images into easily accessible repositories on the Internet is enabling new scientific studies based on the previously unavailable data. Unfortunately, potential scientific advances are hindered by the lack of high-quality and pertinent metadata associated with the image collections. Metadata is required to search the repositories for the imaged specimens needed for a particular study. Since the collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible.  Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.

 Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been acquired worldwide by museums and other research institutions, undertakings that have produced large image collections. 
The metadata that is vital for subsequent machine learning, analysis and
scientific discovery based on these specimen image repositories
is often unavailable, sparse or incorrect.
As a step towards improving metadata in specimen research image collections, our team is developing methods for automatically analyzing fish images to extract a variety of important features. These fish specimens are being studied for a larger project entitled Biology Guided Neural Networks (BGNN), which is developing a novel class of artificial neural networks that can exploit the machine readable and predictive knowledge about biology that is available in specimen images, phylogenies and anatomy ontologies. Using a combination of machine learning and image informatics tools and techniques, we can accurately determine metadata such as fish quantity and location within images, fish orientation and other quantitative fish features, image scaling based on ruler identification and measurement, and general image quality metrics for a substantial number of the images being used in the BGNN project. 
Our goal is to develop image metadata generation methods that both support
the research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing quantitative features in support of specific biological investigations.

% Metadata is often unavailable, sparse or incorrect within specimen image repositories, but is vital for subsequent machine learning, analysis and scientific discovery. Our goal is to develop image metadata generation methods that both support the novel machine learning research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing specific quantitative features in support of specific biological investigations. For example, we are able to automatically measure the length of a fish specimen.  Associating these measurements with location and acquisition date may provide insights into the influence of habitat factors on fish development/health. Since it is prohibitively expensive for scientists to manually gather this data, we are also interested in applying our tools to images of other species stored in a variety of repositories (e.g. iDigBio). The technical challenges in achieving a broader usage of our approach mostly involve training new classifiers for different types of species, learning to segment and read annotation tags, and generalizing our classifier to find and interpret different types of rulers. This presentation will report on our current efforts to automatically generate metadata for fish specimen images and offer thoughts on how to extend these techniques for other specimen image collections.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10002951.10002952.10003219.10003215</concept_id>
       <concept_desc>Information systems~Extraction, transformation and loading</concept_desc>
       <concept_significance>500</concept_significance>
   </concept>
   <concept>
        <concept_id>10010147.10010178</concept_id>
	<concept_desc>Computing methodologies~Artificial intelligence~Computer Vision</concept_desc>
	<concept_significance>500</concept_significance>
    </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Extraction, transformation and loading}
\ccsdesc[500]{Computing methodologies~Artificial intelligence}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{bioinformatics, metadata, image analysis, applied machine learning}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \centering
  \includegraphics[width=.45\textwidth]{images/teaser1_crop}
  \hspace{4mm}
  \includegraphics[width=.45\textwidth]{images/teaser2_crop}
  \caption{Initial object detection on specimen images using Detectron2~\cite{wu2019detectron2}}
  \Description{Fish predictions}
  \label{fig:teaser}
\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

\todo{Sentences I didn't use: Over the last few decades, cyberinfrastructure development, advances in digitization, and the open science tenor, have helped support global access to natural history. OR
Global access to natural history collections has accelerated over the last few decades to advances in cyberinfrastructure, digitization, and the general open science tenor}

Over the last several decades advances in computing, imaging, and
cyberinfrastructure have supported the growth of digital natural history collections; thus supporting the trends towards open scienc.
In the U.S., programs, such as the National Science Foundation’s former
Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for
nearly a decade, and the current Infrastructure Capacity for Biological Research program have supported the digital acquisition and dissemination of
data and images for millions of biological specimens. These resources are globally available through a wide range of portals and digital libraries for researchers, educators, students, and the general public to examine and compare in ways previously unimaginable. Significant challenges surface, however, when researchers seek to apply computational methods to these data items,
due to both image quality and metadata issues.
% (As researchers seek to apply computational methods, and explore AI, significant challenges surface..)

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images, e.g. with low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements, provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata, detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images
to determine if, for instance, a specific fish grows to different lengths
in different habitats.

@@

\section{Related Work}
\textbf{Todo}
\subsection{Metadata for Image Collection}
\todo{Wasn't looking for a citation but stumbled upon this paper: \url{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0207636}, might be worth a skim}
Metadata development is intertwined with the development of digital technology. A wide range of metadata standards are applicable to metadata supporting image description as well as technical aspects, and have been used in a range of image repositories. Examples of Descriptive schemes include X, y, z... Examples of technical schemes include... The use varies across communities, and application technique is primarily manual. Research evaluating such schemes XXX, show XXX.

Natural history collections generally use a sub-set of such standards. Darwin Core, a globally adopted standard overseen by the Taxonomic Data Working Group (TDWG).  [Jane continue here on some of this, and then with preservation/image quality less]

Fish species dataset in Pakistan Shah 2019 \cite{Shah2019FishPakFS},

\subsection{Metadata generation?}

@@

@@

\subsection{Metadata Generation}
Including what has already been discussed, the following metadata properties are being currently being generated:

\begin{table}[t]
    \centering
    \caption{Metadata properties}
    \label{tab:properties}
    \begin{tabular}{cccp{0.5\linewidth}}
        \toprule
        \textbf{Property} & \textbf{Association} & \textbf{Type} & \textbf{Explanation}\\
        \midrule
        \verb|has_fish| & Overall Image & Boolean & Whether a fish was found in the image.\\
        \verb|fish_count| & Overall Image & Integer & The quantity of fish present.\\
        \verb|has_ruler| & Overall Image & Boolean & Whether a ruler was found in the image.\\
        \verb|ruler_bbox| & Overall Image & 4 Tuple & The bounding box of the ruler (if found).\\
        \verb|scale| & Overall Image & Float & The scale of the image in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|bbox| & Per Fish & 4 Tuple & The top left and bottom right coordinates of what define the bounding box for a fish.\\
        \verb|background.mean| & Per Fish & Float & The mean intensity of the background within a given fish's bounding box.\\
        \verb|background.std| & Per Fish & Float & The standard deviation of the background within a given fish's bounding box.\\
        \verb|foreground.mean| & Per Fish & Float & The mean intensity of the foreground within a given fish's bounding box.\\
        \verb|foreground.std| & Per Fish & Float & The standard deviation of the foreground within a given fish's bounding box.\\
        \verb|centroid| & Per Fish & 4 Tuple & The centroid of a given fish's bitmask.\\
        \verb|primary_axis| & Per Fish & 2D Vector & The unit length primary axis (eigenvector) for the bitmask of a given fish.\\
        \verb|clock_value| & Per Fish & Integer & A fish's primary axis converted into an integer ``clock value'' between 1 and 12.\\
        \verb|length| & Per Fish & Float & The length of a fish in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|mask| & Per Fish & 2D Matrix & The bitmask of a fish in 0's and 1's.\\
        \verb|pixel_analysis_failed| & Per Fish & Boolean & Whether the pixel analysis process failed for a given fish. If \verb|true|, \verb|detectron|'s mask and bounding box were used for metadata generation.\\
        \verb|score| & Per Fish & Float & The percent confidence score output by \verb|detectron| for a given fish.\\
        \verb|has_eye| & Per Fish & Boolean & Whether an eye was found for a given fish.\\
        \verb|eye_center| & Per Fish & 2 Tuple & The centroid of a fish's eye.\\
        \verb|side| & Per Fish & String & The side (i.e.\ \verb|'left'| or \verb|'right'|) of the fish that is facing the camera (dependent on finding its eye).\\
      \bottomrule
\end{tabular}
\end{table}

\verb|has_fish|, \verb|fish_count|, \verb|has_ruler|, \verb|ruler_bbox|, \verb|background.*|, \verb|foreground.*|, \verb|bbox|, \verb|mask|, \verb|score|, and \verb|has_eye| have already been discussed. The remaining properties are determined as follows:
\subsubsection{centroid and eye\_center}
Centroids are provided for masks/bounding boxes generated by \verb|detectron|, and since we do not recalculate the mask of fish eyes we can use that value directly for \verb|eye_center|.

Since we recalculate the mask of the fish, we must recalculate its centroid as well. This can be done via
\begin{equation}
    (\bar{x}, \bar{y}) = (\mathrm{round}(\frac{M_{10}}{M_{00}}), \mathrm{round}(\frac{M_{01}}{M_{00}}))
\end{equation}
where \(M_{00}\) is the pixel area of the fish's blob, \(M_{10}\) is the sum of all the \(x\) values of blob pixels, and \(M_{01}\) is the sum of all the \(y\) values of blob pixels.
\subsubsection{side}
Determining which \verb|side| of the fish is visible is predicated on finding its eye. If an eye is found, the sign of the \(x\) component of the vector from the centroid of the fish to the centroid of the eye specifies which side is up: negative for left and positive for right. This assumes the fish was photographed vertically, which is essentially always the case for all image collections our group has worked on.
\subsubsection{primary\_axis and clock\_value}
The \verb|primary_axis| of a fish can be calculated by taking the covariance\todo{I could explain this in more detail if necessary} of its blob in \(x\) and \(y\), which yields its principle eigenvector. This can be directly assigned to the property. If an eye is present, we specify this axis in the direction of the eye in relation to the fish's centroid.

Our team encoded this information as a ``clock value'' between 1 and 12 when recording it by hand. To convert \verb|principal_axis| to \verb|clock_value|, the sign of \(x\) and \(y\) on the principal axis are used to determine which Cartesian quadrant the fish is angles into relative to its centroid. Depending on this quadrant, we then dot product the principal axis with either \([-1,0]\), \([0,-1]\), \([1,0]\) or \([0,1]\) which correspond to 9, 6, 3 and ``0'' o'clock respectively. The resulting radian value is then converted to a polar displacement in clock value space, and added to the comparative clock value used in the dot product. This then gives the fish's clock value from 0 to \(11.\overline{9}\). Before recording as \verb|clock_value| in the output, the value is rounded to the nearest integer, and in the event of a 0 final result it is replaced with 12.
\subsubsection{scale and length}
\(\frac{\mathrm{pixels}}{\mathrm{inch}}\) can be calculated by measuring the distance between the digits 2 and 3 found on the ruler by \verb|detectron|. Converting this to \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) gives the \verb|scale| metadata property as reported in the output.

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction. Since fish are normally measured in a straight line from their snout down the middle of their trunk, we project every pixel onto the major axis of the fish (as a line through its centroid) before checking the distance. This projection is done by intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line. After checking every pixel, measuring the distance between the two furthest projected points gives the length of the fish in pixels. Multiplying this by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{Leipzig2021.01.28.428644} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and meet our current admissibility criteria for what \verb|detectron| and the pixel analysis process can reliably handle. We ran the metadata extraction program on these \(7,244\) images.

@@

\subsection{Fish Detection}
All images in the INHS dataset contain exactly one fish. For \(7,209\) of the specimen images, one fish was detected. For 25 of the images, 2 fish were detected, 3 fish were detected for 3 images, and for 7 images no fish were found. In the case of greater than 1 fish, 9 of the 28 contained tags which overlapped the fish and were themselves labeled as a second fish. Of the remaining 17, \verb|detectron| erroneously labeled the fish as two separate fish objects, or labeled a subsection of the fish a second time.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% REMOVE FOLLOWING LINE BEFORE SUBMITTING %%%%%%%%%
\setlength {\marginparwidth }{2cm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{todonotes}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{xurl}
\usepackage{booktabs}
\usepackage{float}
\usepackage[noadjust]{cite}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Automated Metadata Generation for Fish Specimen Image Collections%*\\
    {\footnotesize \thanks{NSF Grant ???} }}
%do we add grant now, or wait til acceptance. i think we need to ultimately remove any identifying info., i think bgnn is likely okay; but, perhaps mention metadata research center (MRC) need to be removed?. if so, i kind of like to put it back in for final version, if accepted. it helps with getting funding ... you are both listed on the MRC page. we should check on the identifying info. removed, or not.
%for the record: Research supported by NSF OAC #1940233 and #1940322.

\author{
\IEEEauthorblockN{Joel Pepper}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Drexel University}\\
Philadelphia, USA\\
0000-0002-1601-8729}
\and
\IEEEauthorblockN{Jane Greenberg}
\IEEEauthorblockA{\textit{Department of Information Science} \\
\textit{Drexel University}\\
Philadelphia, USA\\
0000-0001-7819-5360}
\and
\IEEEauthorblockN{Yasin Baki\c{s}}
\IEEEauthorblockA{\textit{Biodiversity Research Institute} \\
\textit{Tulane University}\\
New Orleans, USA\\
0000-0001-6144-9440}
\and
\IEEEauthorblockN{Xiaojun Wang}
\IEEEauthorblockA{\textit{Biodiversity Research Institute} \\
\textit{Tulane University}\\
New Orleans, USA\\
0000-0001-8639-6795}
\and
\IEEEauthorblockN{Henry Bart Jr.}
\IEEEauthorblockA{\textit{Biodiversity Research Institute} \\
\textit{Tulane University}\\
New Orleans, USA\\
0000-0002-5662-9444}
\and
\IEEEauthorblockN{David Breen}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Drexel University}\\
Philadelphia, USA\\
0000-0002-1376-5008}
}

\maketitle

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
     Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been acquired worldwide by museums and other research institutions, undertakings that have produced large image collections.
The metadata that is vital for subsequent machine learning, analysis and
scientific discovery based on these specimen image repositories
is often unavailable, sparse or incorrect.
As a step towards improving metadata in specimen research image collections, our team is developing methods for automatically analyzing fish images to extract a variety of important features. These fish specimens are being studied for a larger project entitled Biology Guided Neural Networks (BGNN), which is developing a novel class of artificial neural networks that can exploit the machine readable and predictive knowledge about biology that is available in specimen images, phylogenies and anatomy ontologies. Using a combination of machine learning and image informatics tools and techniques, we can accurately determine metadata such as fish quantity and location within images, fish orientation and other quantitative fish features, image scaling based on ruler identification and measurement, and general image quality metrics for a substantial number of the images being used in the BGNN project.
Our goal is to develop image metadata generation methods that both support
the research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing quantitative features in support of specific biological investigations.
\end{abstract}

\begin{IEEEkeywords}
bioinformatics, metadata, image analysis, applied machine learning
\end{IEEEkeywords}

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which have specimen images.~\cite{beaman2012mass} Additionally, initiatives, such as National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for nearly a decade, have supported the digitization and curation of tens of thousands of biological specimens.~\cite{page2015digitization} These digital specimen accessible through global, open access repositories; they enable researchers, educators, students, and the general public to examine and compare in ways previously unimaginable. Significant challenges surface, however, when researchers seek to apply computational methods to these data items, due to both image quality and metadata issues.
%{jane}should we try to insert machine learning here, last sentence, or computation methods ok?

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images, e.g. with low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements, provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata,~\cite{biodiv_info_standards} detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images
to determine if, for instance, a specific fish grows to different lengths
in different habitats.

@@

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. Darwin Core~\cite{biodiv_info_standards} developed specifically for digital museum projects is one of the most frequently used standards. Another approach is to create a metadata application profile by integrating properties from Dublin Core's DCMI Metadata Terms \cite{dc_terms}, the Ecological Metadata Language (EML) \cite{EML_2019} and other namespaces. There is also the recent Audubon Core supporting the discoverability, dissemination, and use of data relating to biological organisms, including digitized specimens.\cite{audub_core} These descriptive standards include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the eye of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is limited, particularly for researchers seeking to apply machine learning, and leverage both the image and the metadata for scientific research.

This limitation is magnified when trying get a handle on the actual quality of the digitized specimen. Descriptive oriented standards support search and retrieval, and the biodiversity community advocated for data fitness standards \cite{chapman2020developing}. Even so, metadata properties used to record image quality is severely limited and generally missing. This point is addressed in detail in \cite{leipzig2021biodiversity}, justifying the capture of 22 metadata properties supporting selection of high quality images for the development of a bgnn. Although, the manual generation of this metadata is costly in terms of labor and time.

%Fish species dataset in Pakistan Shah 2019 \cite{Shah2019FishPakFS},

\subsection{Automatic Metadata Generation}
Technical metadata, such as camera settings and temporal information (date and time) are automatically generated during a digitization sequence, following standards such as XX. Some of this metadata may be useful in for selecting a machine learning sample. A researcher may desire images captured chiefly with a certain aperture setting. Even so, the majority of automatically captured technical metadata associated with digitized specimens is insufficient for computational research, leaving researchers to reply on manually generated descriptive metadata, which itself is sparse.

Automatic metadata generation of descriptive bibliographic data has been a research focus for close to 20 years \cite{liddy2002automatic,greenberg2004metadata,cardinaels2005automating}. Researchers have applied support vector machine (SVM) approaches \cite{han2003automatic}, and associated networks to address sparse and incomplete metadata \cite{rodriguez2009automatic}, and various successes are integrated into day-to-day workflows.

%\cite{heidorn2008automatic} speciment labels \cite{manso2004automatic} cartographic info.
%above paragraph..need to finesse a couple of sentences w/results, weave in these 2 citations, and then say we can look to image analysis.

@@

\begin{table*}[t]
    \centering
    \caption{Metadata properties}
    \label{tab:properties}
    \begin{tabular}{cccp{0.5\linewidth}}
        \toprule
        \textbf{Property} & \textbf{Association} & \textbf{Type} & \textbf{Explanation}\\
        \midrule
        \verb|has_fish| & Overall Image & Boolean & Whether a fish was found in the image.\\
        \verb|fish_count| & Overall Image & Integer & The quantity of fish present.\\
        \verb|has_ruler| & Overall Image & Boolean & Whether a ruler was found in the image.\\
        \verb|ruler_bbox| & Overall Image & 4 Tuple & The bounding box of the ruler (if found).\\
        \verb|scale| & Overall Image & Float & The scale of the image in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|bbox| & Per Fish & 4 Tuple & The top left and bottom right coordinates of the bounding box for a fish.\\
        \verb|background.mean| & Per Fish & Float & The mean intensity of the background within a given fish's bounding box.\\
        \verb|background.std| & Per Fish & Float & The standard deviation of the background within a given fish's bounding box.\\
        \verb|foreground.mean| & Per Fish & Float & The mean intensity of the foreground within a given fish's bounding box.\\
        \verb|foreground.std| & Per Fish & Float & The standard deviation of the foreground within a given fish's bounding box.\\
        \verb|centroid| & Per Fish & 4 Tuple & The centroid of a given fish's bitmask.\\
        \verb|primary_axis| & Per Fish & 2D Vector & The unit length primary axis (eigenvector) for the bitmask of a given fish.\\
        \verb|clock_value| & Per Fish & Integer & Fish's primary axis converted into an integer ``clock value'' between 1 and 12.\\
        \verb|length| & Per Fish & Float & The length of a fish in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|mask| & Per Fish & 2D Matrix & The bitmask of a fish in 0's and 1's.\\
        \verb|pixel_analysis_failed| & Per Fish & Boolean & Whether the pixel analysis process failed for a given fish. If \verb|true|, \verb|detectron|'s mask and bounding box were used for metadata generation.\\
        \verb|score| & Per Fish & Float & The percent confidence score output by \verb|detectron| for a given fish.\\
        \verb|has_eye| & Per Fish & Boolean & Whether an eye was found for a given fish.\\
        \verb|eye_center| & Per Fish & 2 Tuple & The centroid of a fish's eye.\\
        \verb|side| & Per Fish & String & The side (i.e.\ \verb|'left'| or \verb|'right'|) of the fish that is facing the camera (dependent on finding its eye).\\
      \bottomrule
\end{tabular}
\end{table*}

@@

\subsection{Metadata Generation}

The following metadata properties are generated from the methods
described above:
\verb|has_fish|, \verb|fish_count|, \verb|has_ruler|, \verb|ruler_bbox|, \verb|background.{mean,std}|, \verb|foreground.{mean,std}|, \verb|bbox|, \verb|mask|, \verb|score|, and \verb|has_eye|.
The remaining properties are computed as described below, with all the
metadata properties listed in Table \ref{tab:properties}.

\subsubsection{centroid and eye\_center}
Centroids are provided for the masks and bounding boxes generated by \verb|detectron|, and since we do not recalculate the mask of fish eyes we can use that value directly for \verb|eye_center|.

Since we recalculate the mask of the fish, we must recalculate its centroid as well. This can be done via
\begin{equation}
    (\bar{x}, \bar{y}) = (\mathrm{round}(\frac{M_{10}}{M_{00}}), \mathrm{round}(\frac{M_{01}}{M_{00}})),
\end{equation}
where \(M_{00}\) is the pixel area of the fish's blob, \(M_{10}\) is the sum of all the \(x\) values of blob pixels, and \(M_{01}\) is the sum of all the \(y\) values of blob pixels.
\subsubsection{side}
Determining which \verb|side| of the fish is visible is predicated on finding its eye. If an eye is found, the sign of the \(x\) component of the vector from the centroid of the fish to the centroid of the eye specifies which side is up: negative for left and positive for right. This assumes the fish was photographed vertically, which is essentially always the case for all image collections our group has worked on.
\subsubsection{primary\_axis and clock\_value}
The \verb|primary_axis| of a fish can be calculated by taking the covariance\todo{I could explain this in more detail if necessary. (DEB No, that's OK.} of its blob in \(x\) and \(y\), which yields its principle eigenvector.
The eigenvector can be directly assigned to the property. If an eye is
present, we ensure that \verb|primary_axis| points in the direction of the
eye in relation to the fish's centroid.

Our team encoded this information as a ``clock value'' between 1 and 12 when manually recording it. To convert \verb|principal_axis| to \verb|clock_value|, the sign of \(x\) and \(y\) on the principal axis are used to determine which Cartesian quadrant the fish angles into relative to its centroid. 
Depending on this quadrant, we dot product the principal axis with either \([-1,0]\), \([0,-1]\), \([1,0]\) or \([0,1]\), which correspond to 9, 6, 3 and ``0'' o'clock respectively. The resulting radian value is then converted to a polar displacement in clock value space, and added to the comparative clock value used in the dot product. This gives the fish's clock value from
0 to \(11.\overline{9}\). Before recording \verb|clock_value| in the output,
the value is rounded to the nearest integer, with a 0 final result replaced with 12.
\subsubsection{scale and length}
\(\frac{\mathrm{pixels}}{\mathrm{inch}}\) can be calculated by measuring the distance in pixels between the digits 2 and 3 (a 1 inch separation) found on the ruler by \verb|detectron|. Converting this to \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) gives the \verb|scale| metadata property as reported in the output.

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{Leipzig2021.01.28.428644} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel analysis processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 50 specimens from the set of \(7,244\) were analyzed by
hand for comparison.

@@

\subsection{Fish Detection}
All images in the INHS dataset contain exactly one fish. For \(7,209\) of the specimen images, one fish was detected, a 99.5\% correct rate.
For 25 of the images, 2 fish were detected, 3 fish were detected for 3 images, and for 7 images no fish were found. In the case of greater than 1 fish, 9 of the 28 contained tags which overlapped the fish and were themselves labeled as a second fish. Of the remaining 17, \verb|detectron| erroneously labeled the fish as two separate fish objects, or labeled a subsection of the fish a second time.
